Symmetric Circuits for Rank Gates

### 2018-02-05 sections/symmetric-circuits.tex
February 5 2018

% AD - rewrote the paragraph to reduce the use of subscripts and superscripts.
Let $\tau := (R, S, \nu)$ be a many-sorted vocabulary. Let $D := \biguplus_{s \in
  S} D_{s} = \{(s,d) : d \in D_s\}$, be a disjoint union of non-empty sets,
indexed by $S$. For $R_i \in R$ let $R^D_i$ denote the trivial relation over the
many-sorted universe $D$, i.e.\ $R^D_i := D_{a_1} \times \ldots \times
D_{a_{r}}$, where $\nu(R_i) = (a_1, \ldots , a_{r)})$. There is an obvious
interpretation of $\tau$ over $D$ given by assigning each sort symbol in $\tau$
in the obvious manner and each relation symbol in $\tau$ to the trivial relation
over $D$. We call this \emph{the structure defined by $(\tau, D)$}, and denote
it by $\str{\tau, D}$.

@@

We abbreviate $(\vec{a}, R) \in \ind(\tau, D)$ by $\vec{a}_R$, and refer
to $R$ as \emph{the type} of $\vec{a}_R$. We often abuse notation and treat
$\vec{a}_R$ as the tuple $\vec{a}$ in $R^D$. In the event that there is only one
relational symbol in $\tau$ we identify $\vec{a}_R$ with $\vec{a}$.

@@

% GW: move sorted permutations into the background section
For the remainder of this subsection we fix a vocabulary $\tau$ and a
many-sorted set $D := \biguplus_{s \in S} D_{s}$, let $X := \ind (\tau, D)$ and $G \leq \sym_{X}$.
There is an obvious action of $G$ on $X$. There is also an action of the
automorphism group $\str{\tau, D}$ on $X$ defined for the automorphism $\sigma$
by $\sigma \cdot \vec{a}_R := (\sigma \vec{a})_R$. We denote the group of
automorphisms of $\str{\tau, D}$ by $\aut(\tau, D)$, and identify it with the
subgroup of $\sym_X$ that acts equivalently on $X$.

We can extend the action of $G$ on $X$ to an action on the set of functions of
the form $f: X \rightarrow H$, for any set $H$, defined by $\sigma \cdot f(x) :=
f(\sigma \cdot x)$ for all $x \in X$, $\sigma \in G$.

@@

We say that two functions $f: \ind(\tau, D) \rightarrow H$ and $g: \ind(\tau,
D') \rightarrow H$ are \emph{isomorphism-equivalent} if there exists an
isomorphism $\pi :\str{\tau, D} \rightarrow \str{\tau, D'}$ such that $f (\vec{a}_R) =
g (\pi \vec{a}_R)$ for all $\vec{a}_R \in \ind(\tau, D)$. It is easy to see that
isomorphism-equivalence is an equivalence relation.

We are often interested in the case where $X = \ind(\tau, D) = \ind(\tau, D')$.
We say that $f, g: X \rightarrow H$ are \emph{$G$-equivalent} if there exists
$\sigma \in G$ such that $f = \sigma \cdot g$. We note that in this case $f$ and
$g$ are isomorphism-equivalent if, and only if, they are $\aut(\tau,
D)$-equivalent.

In this paper we will often take the co-domain $H$ to be $\{0,1\}$, and so $f :
X \rightarrow \{0,1\}$. In this case we may think of $f$ as defining for each
relation $R^D_i$ a subset of that relation, and so determining a
$\tau$-structure over the universe $D$. We note that two functions $f, g: X
\rightarrow \{0,1\}$ are isomorphism-equivalent if, and only if, the two
$\tau$-structures they determine are isomorphic.

@@

In each case a natural choice of $\tau$ (and $D$) indexes the input to the
structured function appropriately and produces the appropriate symmetry
condition. It is worth mentioning that the graph-invariant case is not strictly
reproduced using structured functions. This is because structured functions
index the input string using the elements of the relations of the complete
$\tau$-structure (in this case, edges in the complete graph). As such, the input
string is indexed by ordered pairs of elements, rather than two-element
subsets, and as such the symmetry condition on the relation is not enforced. As
such, in the structured function case the function decides properties of
directed graphs, rather than ordinary graphs. We should note, however, that the symmetry
of the edge relation is a property of directed-graph, and as such can be decided
by an isomorphism-invariant structured function over the graph vocabulary.

% , and any
% input sequence encodes a sub-relation of that complete relation. As such, the
% function  This is because in the event that $\tau := {E}$ (where $E$ is
% a binary relation, thought of as the edge relation) there is no assurance that
% the function $f : \ind
% (\tau, D) \rightarrow \{0,1\}$ is encoding

% as the structured functions are rather defined over directed
% graphs.

@@

In keeping with convention for Boolean functions, and for introducing simplicity
without a loss of generality, we assume that the universe of any structured
function is always a disjoint union of initial segments of the natural numbers,
i.e.\ if $F$ is structured function with universe $D_{s_1} \uplus \ldots \uplus
D_{s_p}$, then for all $s \in \{s_1 , \ldots , s_p\}$ there exists $d_s \in
\nats{N}$ such that $D_{s} = [d_s]$. We also assume that every $(\tau,
D)$-structured function has the property that every $a \in D$ appears in at
least one element of $\ind(\tau, D)$.

@@

We now define a \emph{$(\mathbb{B}, \rho)$-circuit} . This is a circuit that
computes a function on structures in the relational vocabulary $\rho$ and where
the gates are labelled by elements of the basis $\mathbb{B}$ consisting of
structured functions. Note that, in the interests of keeping our definitions
from getting excessively complicated, we do not consider circuits computing
functions over multi-sorted vocabularies.

@@

Let $C = \langle G, \Omega, \Sigma, \Lambda, L \rangle$ be a $(\mathbb{B},
\rho)$-circuit of order $n$. The order of a circuit denotes the fact that it
takes as input structures of size $n$. We often consider families of circuits of
the form $(C_n)_{n \in \nats}$, and here we assume that each $C_n$ is a circuit
of order $n$. For a gate $g \in C$, we define the \emph{index} of $g$, denoted
by $\ind(g)$, as $\ind(\Sigma(g))$. We define the relation $W \subseteq G^2$ by
$W(h,g)$ if, and only if, $h \in \range(L(g))$. For each $g \in G$ we let
$W(\cdot, g) := \{h \in G : W(h,g)\}$ and $W(g, \cdot) := \{h \in G : W(g,h)\}$.
We call the elements of $W(\cdot, g)$ the \emph{children} of $g$ and the
elements of $W(g, \cdot)$ the \emph{parents} of $g$. We also abbreviate $W(g,
\cdot)$ by $H_g$.

@@

  We now present a number of observations, and then combine these observations
  to prove both directions of the hypothesis. Let $\mathcal{B}$ be a
  $\rho$-structure of size $n$ over the universe $[n]$. We can define a function
  $f_{\mathcal{B}} : \ind(\rho, [n]) \rightarrow {0,1}$ such that
  $f_{\mathcal{B}} (\vec{a}_R) = 1$ if, and only if, $\vec{a} \in
  R^{\mathcal{B}}$. Let $f : \ind(\rho, [n]) \rightarrow \{0,1\}$. We can define
  the $\rho$-structure $\mathcal{B}_f$ over the universe $[n]$ such that for all
  $R \in \rho$, we have $\vec{a} \in R^{\mathcal{B}_f}$ if, and only if,
  $f(\vec{a}_R) = 1$. It is easy to see that the functions $\mathcal{B} \mapsto
  f_{\mathcal{B}}$ and $f \mapsto \mathcal{B}_f$ are inverse to one another, and
  so define a bijection. Moreover, we notice that for all $\sigma \in \sym_n$ we
  have that $f_{\sigma \mathcal{B}}(\vec{a}_R) = 1$ if, and only if, $\vec{a}
  \in R^{\sigma \mathcal{B}}$ if, and only if, $\sigma^{-1} \vec{a} \in
  R^{\mathcal{B}}$ if, and only if, $\sigma ^{-1} f_{\mathcal{B}}(\vec{a}_R) =
  f_{\mathcal{B}}(\sigma^{-1} \vec{a}_R)= 1$.

@@

  Let $\gamma_1, \gamma_2 \in [n]^{\underline{U}}$ and let $\sigma \in \sym_n$
  be such that $\gamma_1 = \sigma \gamma_2$. We have that
  \begin{align*}
    C[\gamma_1 \mathcal{A}](g_{\text{o}}) = C[\gamma_2 \mathcal{A}](g_{\text{o}}) & \Leftrightarrow  F_C(f_{\gamma_1 \mathcal{A}}) = F_C (f_{\gamma_2 \mathcal{A}}) \\
                                                                                  & \Leftrightarrow F_C(f_{\gamma_1 \mathcal{A}}) = F_C (f_{\sigma^{-1}\gamma_1 \mathcal{A}}) \\
                                                                                  & \Leftrightarrow F_C(f_{\gamma_1 \mathcal{A}}) = F_C (\sigma f_{\gamma_1 \mathcal{A}}).
  \end{align*}

  We now combine the above observations to prove the result. Suppose $F_C$ is
  isomorphism-invariant. Let $\gamma_1, \gamma_2 \in [n]^{\underline{U}}$ and
  let $\sigma \in \sym_n$ be such that $\gamma_1 = \sigma \gamma_2$. Then, from
  isomorphism-invariance, we have $F_C(f_{\gamma_1 \mathcal{A}}) = F_C (\sigma
  f_{\gamma_1 \mathcal{A}})$, and so $C[\gamma_1 \mathcal{A}](g_{\text{o}}) =
  C[\gamma_2 \mathcal{A}](g_{\text{o}})$.

  Suppose $C$ is invariant. Fix the bijection $\gamma_1 \in
  [n]^{\underline{U}}$. Let $\sigma \in \sym_n$ and $f: \ind(\rho, [n])
  \rightarrow \{0,1\}$. Let $\mathcal{A} = \gamma^{-1}_1 \mathcal{B}_f$ and
  $\gamma_2 := \sigma^{-1}\gamma_1 $. Since $C$ is invariant we have $C[\gamma_1
  \mathcal{A}](g_{\text{o}}) = C[\gamma_2\mathcal{A}](g_{\text{o}})$, and so
  $F_C (\sigma f_{\gamma_1 \mathcal{A}}) = F_C(f_{\gamma_1 \mathcal{A}})$. But
  $f_{\gamma_1 \mathcal{A}} = f_{\mathcal{B}_f} = f$, and so $F_C (\sigma f) =
  F_C (f)$. It follows that $F_C$ is isomorphism-invariant.

@@

It follows that for any symmetric circuit $C$ of order $n$ there is a
homomorphism $h$ that maps $\sym_n$ to $\aut(C)$, taking any $\sigma \in \sym_n$
to an automorphism that extends it. Suppose $C$ contains no relational gates of
non-zero arity. In that case $C$ computes a constant function. As such, in this
paper we always assume a circuit contains at least one relational gate with
non-zero arity. There there exists a relational gate in $C$ with some element of
$[n]$ appearing in the tuple labelling that gate. By symmetry it follows that
all elements of $[n]$ appear in tuples labelling relational gates in $C$. Thus
no two distinct elements of $\sym_n$ agree on all input gates, and so the
homomorphism $h$ is injective.

@@

Anderson and Dawar's argument makes use of the fact that many important circuit
properties are polynomial-time decidable for circuits defined over a symmetric
bases. We show in Section~\label{sec:transparency-and-unique-extension} that for
circuits that include gates computing non-symmetric functions most of these
properties are at least as hard to decide as the graph-isomorphism problem. As
such, constructing an analogous translation, as well as establishing the many
other results that rely on certain properties being polynomial-time decidable,
would be highly non-trivial.

In order to deal with this problem we limit the ways in which gates computing
non-symmetric functions can be wired into the circuit and we show in
Section~\label{sec:transparency-and-unique-extensions} that this limitation is
sufficient for our purposes.

% ourselves to circuits in which those gates that compute non-symmetric
% functions

% While we should like to establish an analogous polynomial-time computable
% translation from a general circuit to a equivalent circuit with unique
% extensions, it turns out that the inclusion of gates that compute
% non-symmetric functions greatly complicates the analysis of the circuit. We
% show in Section~\label{sec:transparency-and-unique-extensions} that most of
% the polynomial-time decidable properties used by Anderson and Dawar in order
% to define this translation are at least as hard to decide for general circuits
% as the graph-isomorphism problem.

% Since the proof of our main result requires that many circuit properties are
% polynomial-time decidable, we introduce restrictions on how gates computing
% non-symmetric functions can be wired into the circuit.

% The inclusion of gates that inclusion greatly complicates the analysis of a
% general circuit, resulting in many circuit polynomial-time decidable circuit
% properties

% It turns out that allowing non-symmetric functions greatly complicates the the
% structure of a circuit, and makes certain many circuit properties difficult to
% decide (in a technical sense).

% inclusion of non-symmetric functions in the basis makes deciding certain
% circuits properties non-trivial,

% determining certain properties

% in the basis greatly complicates the argument, unfortunately, for the general
% $(\mathbb{B}, \rho)$-circuit, establishing an exactly analogous condition is a
% non-trivial problem. We discuss this problem in more detail in
% Section \label{sec:transparency-and-unique-extensions}
.

We now introduce the notion of a circuit with \emph{unique labels}, and show
that if a circuit has unique labels then it has unique extensions. Before we do
so, we define the \emph{syntactic equivalence} relation on the gates of a
circuit.

@@

We could define an equivalence relation on gates that identifies a pair of gates
if, and only if, they compute the same function in the circuit. However, it is
not hard to show that deciding this equivalence relation is $\NP$-hard. The
syntactic-equivalence relation instead identifies gates that have `hereditary
similar' circuits underneath them. Since it can be shown that two
syntactically-equivalent gates must compute the same function in the circuit, it
follows that the syntactic-equivalence relation refines the above-referenced
equivalence relation.

We may also define an equivalence relation between gates identifying a pair of
gates $(h, h')$ in a circuit $C$ of order $n$ if, and only if, there exists
$\sigma \in \sym_n$, $\pi, \pi' \in \aut(C)$ both extending $\sigma$, and $g$ a
gate in $C$, such that $h = \pi (g)$ and $h' = \pi' (g)$. We now show this
relation is a refinement of syntactic-equivalence.

@@

It is easy to show that, for a circuit $C$ of order $n$, if $\sigma \in \sym_n$
and both $\pi, \pi' \in \aut(C)$ extend $\sigma$, then for every gate $g$ in
$C$, $\pi (g)$ and $\pi'(g)$ are syntactically-equivalent.

We now introduce the \emph{unique labels} and \emph{transparency} conditions on
circuits.

@@

There is a small technical point that needs some discussion before we continue.
The definition of a circuit allows for the inclusion of gates whose outputs need
not lead to any output gate. In other words, we can be sure, not matter the rest
of the circuit, that these gates cannot have any effect the function computed by
the circuit. We say these gates are \emph{redundant}.

@@

It is easy to see that we can simply remove the redundant gates from a circuit,
and so construct from a given circuit an equivalent circuit without redundant
gates. Let $C := \langle G, \Omega, \Sigma, \Lambda, L \rangle$ be a circuit of
order $n$ and let $g$ be a redundant circuit in $C$. It is easy to see that $g$
is not the child of a non-redundant gate, nor is it a relational or output gate.
Thus we can define $C' := \langle G', \Omega, \restr{\Sigma}{G'}, \Lambda,
\restr{L}{G'} \rangle$, where $G' \subseteq G$ is the set of all non-redundant
gates in $C$.

It is easy to see that $C'$ computes the same query as $C$, and that each $g \in
G'$ has unique labels in $C$ if, and only if, $g$ has unique labels in $C'$.
Moreover, it can be shown that if $C$ is symmetric and $g'$ is in the orbit of
$g$ in $C$ then $g' \in G'$. As such, if $C$ is symmetric then so is $C'$. We
thus have the following Lemma.

\begin{lem}
  Let $C := \langle G, \Omega, \Sigma, \Lambda, L \rangle$ be a circuit of order
  $n$. Let $G'$ be the set of non-redundant gates in $C$. Let $C' := \langle G',
  \Omega, \restr{\Sigma}{G'}, \Lambda, \restr{L}{G'} \rangle$. Then (i) $C$ and
  $C'$ compute the same query, (ii) no gate in $C'$ is redundant, (iii) if $C$
  is symmetric then $C'$ is symmetric, (iv) if $C$ has unique labels then $C'$
  has unique labels, and (v) if $C$ is transparent then $C'$ is transparent
\end{lem}

We note that the construction of $C'$ from $C$ in Lemma~\ref{lem:redudant-gates}
can be implemented so as to run in time polynomial in the size of $C$. As such,
for the remainder of this paper we assume every circuit has no redundant gates.
With this assumption stated we now show that every circuit with unique labels
has unique extensions.

@@

\begin{prop}
  Let $C := \langle G, \Omega, \Sigma, \Lambda, L\rangle$ be a $(\mathbb{B},
  \rho)$-circuit of order $n$. If $C$ has unique labels then $C$ has unique
  extensions.
  \label{prop:unique-labels-unique-extensions}
\end{prop}
% \begin{prop}
%   Let $C_n := \langle G, \Omega, \Sigma, \Lambda, L\rangle$ be an injective
%   $(\mathbb{B}, \tau)$-circuit. Then $C_n$ has unique labels if, and only if,
%   $C_n$ has unique extensions.
% \end{prop}
\begin{proof}
  Suppose $C$ has unique labels. Let $\pi_{\sigma}, \pi_{\sigma}' \in \aut(C_n)$
  be automorphisms extending $\sigma \in \sym_n$. We now prove that $\pi_\sigma
  = \pi_{\sigma}'$. Let $\pi := \pi_{\sigma}'\pi^{-1}_\sigma$. We have that
  $\pi$ is an automorphism of the circuit and that $\pi$ extends
  $\sigma^{-1}\sigma = e$, the identity permutation. It follows that $\pi$ fixes
  all input gates and, since for any output gate $g$, $\pi (g) = \pi
  (\Omega(\vec{a})) = \Omega(e \vec{a}) = \Omega(\vec{a}) = g$ for all $\vec{a}
  \in [n]^q$, $\pi$ fixes all output gates.

@@

We also show that a transparent circuit can be converted into an equivalent
circuit with unique labels in time polynomial in the size of the circuit. Since
we are working with $P$-uniform families of transparent circuits, this fact
allows us to assume, without a loss of generality, that transparent circuits
have unique labels. We prove this formally in Proposition
\ref{prop:transparent-unique}.

@@

In many contexts, circuits are defined over a basis of symbols, usually the
usual logical symbols $\land, \neg, \lor$ and perhaps $\maj$ (for majority).
When evaluating a circuit, a gate labelled by a particular symbol is evaluated
using a Boolean function appropriate to the particular symbol. We have instead
defined a basis to be a set of Boolean functions. However, when reasoning about
circuits or dealing with them as inputs to algorithms we often want to identify
a function in the basis with a particular symbol. We now formally define a few
natural bases used in this paper, assigning to each function in these bases an
appropriate symbol.

>>>>

% AD - rewrote the paragraph to reduce the use of subscripts and superscripts.
Let $\tau := (R, S, \nu)$ be a many-sorted vocabulary. Let $D := \biguplus_{s
  \in S} D_{s} = \{(s,d) : d \in D_s\}$, be a disjoint union of non-empty sets,
indexed by $S$. For $R_i \in R$ let $R^D_i$ denote the trivial relation over the
many-sorted universe $D$, i.e.\ $R^D_i := D_{a_1} \times \ldots \times
D_{a_{r}}$, where $\nu(R_i) = (a_1, \ldots , a_{r)})$. There is an obvious
interpretation of $\tau$ over $D$ given by assigning each sort symbol $s$ in
$\tau$ to the set $D_s$ and each relation symbol in $\tau$ to the trivial
relation over $D$. We call this \emph{the structure defined by $(\tau, D)$}, and
denote it by $\str{\tau, D}$.

@@

We abbreviate $(\vec{a}, R) \in \ind(\tau, D)$ by $\vec{a}_R$, and refer to $R$
as the \emph{tag} of $\vec{a}_R$. We often abuse notation and treat $\vec{a}_R$
as the tuple $\vec{a}$ in $R^D$. In the event that there is only one relational
symbol in $\tau$ we identify $\vec{a}_R$ with $\vec{a}$.

@@

% GW: move sorted permutations into the background section
For the remainder of this subsection we fix a vocabulary $\tau$ and a
many-sorted set $D := \biguplus_{s \in S} D_{s}$. Let $X := \ind (\tau, D)$ and
$G \leq \sym_{X}$. We note that $G$ there is an obvious group action of $G$ on
$X$. Let $\aut (\tau, D)$ be the automorphism group of $\str{\tau, D}$. We
define a group action of $\aut (\tau, D)$ on $X$ for $\sigma \in \aut(\tau, D)$
by $\sigma \cdot \vec{a}_R := (\sigma \vec{a})_R$, for all $\vec{a}_R \in X$. We
identify $\aut(\tau, D)$ with the subgroup of $\sym_X$ that acts equivalently on
$X$.

We can extend the action of $G$ on $X$ to a right group action on the set of
functions of the form $f: X \rightarrow H$, for any set $H$, defined by $(f \cdot
\sigma)(x) := f(\sigma x)$ for all $x \in X$, $\sigma \in G$.

@@

We say that two functions $f: \ind(\tau, D) \rightarrow H$ and $g: \ind(\tau,
D') \rightarrow H$ are \emph{isomorphism-equivalent} if there exists an
isomorphism $\pi :\str{\tau, D} \rightarrow \str{\tau, D'}$ such that
$f(\vec{a}_R) = g(\pi \vec{a}_R)$ for all $\vec{a}_R \in \ind(\tau, D)$. It is
easy to see that isomorphism-equivalence is an equivalence relation.

We are often interested in the case where $X = \ind(\tau, D) = \ind(\tau, D')$.
We say that $f, g: X \rightarrow H$ are \emph{$G$-equivalent} if there exists
$\sigma \in G$ such that $f = g \sigma$. We note that in this case $f$ and $g$
are isomorphism-equivalent if, and only if, they are $\aut(\tau, D)$-equivalent.

In this paper we will often take the co-domain to be $\{0,1\}$, and so have a
function of the form $f : X \rightarrow \{0,1\}$. In this case we may think of
$f$ as defining for each relation $R^D_i$ a subset of that relation, and thus
determining a $\tau$-structure over the universe $D$. We note that two functions
$f, g: X \rightarrow \{0,1\}$ are isomorphism-equivalent if, and only if, the
two $\tau$-structures they determine are isomorphic.

@@

In each case a natural choice of $\tau$ (and $D$) indexes the input to the
structured function appropriately and produces the appropriate symmetry
condition. It is worth mentioning that the graph-invariant case is not strictly
reproduced using structured functions. This is because in the structured
function case the input string is indexed by all possible elements of the binary
relation (i.e. all ordered pairs over the universe). As such, the input string
of the structured function may encode any structure over the universe with a
single binary relation, i.e. a directed graph. In contrast, the graph-invariant
function discussed at the beginning of this subsection indexes the input string
using two element subsets of the universe, rather pairs of elements, an encoding
that naturally enforces the symmetry and irreflexivity properties the edge
relation. We should note, however, that symmetry and irreflexivity are
properties of directed-graph, and as such can be decided by an
isomorphism-invariant structured function over the graph vocabulary.

% , and any input sequence encodes a sub-relation of that complete relation. As
% such, the function This is because in the event that $\tau := {E}$ (where $E$
% is a binary relation, thought of as the edge relation) there is no assurance
% that the function $f : \ind (\tau, D) \rightarrow \{0,1\}$ is encoding

% as the structured functions are rather defined over directed graphs.

@@

In keeping with convention for Boolean functions, and for introducing simplicity
without a loss of generality, we assume that the universe of any structured
function is always a disjoint union of initial segments of the natural numbers,
i.e.\ if $F$ is structured function with universe $D_{s_1} \uplus \ldots \uplus
D_{s_p}$, then for all $s \in \{s_1 , \ldots , s_p\}$ there exists $d_s \in
\nats$ such that $D_{s} = [d_s]$. We also assume that every $(\tau,
D)$-structured function has the property that every $a \in D$ appears in at
least one element of $\ind(\tau, D)$.

@@

We now define a \emph{$(\mathbb{B}, \rho)$-circuit}. This is a circuit that
computes a function on structures in the relational vocabulary $\rho$ and where
the gates are labelled by elements of the basis $\mathbb{B}$ consisting of
structured functions. Note that, in the interests of keeping our definitions
from getting excessively complicated, we do not consider circuits computing
functions over multi-sorted vocabularies.

@@

Let $C = \langle G, \Omega, \Sigma, \Lambda, L \rangle$ be a $(\mathbb{B},
\rho)$-circuit of order $n$. The order of a circuit denotes the fact that it
takes as input structures of size $n$. If $(C_n)_{n \in \nats}$ is a family of
circuits we assume that each $C_n$ is a circuit of order $n$.

For a gate $g \in C$, we let the \emph{index} of $g$, denoted by $\ind(g)$, be
$\ind(\Sigma(g))$. We define the relation $W \subseteq G^2$ by $W(h,g)$ if, and
only if, $h \in \img(L(g))$. For each $g \in G$ we let $W(\cdot, g) := \{h \in G
: W(h,g)\}$ and $W(g, \cdot) := \{h \in G : W(g,h)\}$. We call the elements of
$W(\cdot, g)$ the \emph{children} of $g$ and the elements of $W(g, \cdot)$ the
\emph{parents} of $g$. We also abbreviate $W(g, \cdot)$ by $H_g$.

@@

  We now present a number of observations, and then combine these observations
  to prove both directions of the hypothesis. Let $\mathcal{B}$ be a
  $\rho$-structure of size $n$ over the universe $[n]$. We can define a function
  $f_{\mathcal{B}} : \ind(\rho, [n]) \rightarrow {0,1}$ such that
  $f_{\mathcal{B}} (\vec{a}_R) = 1$ if, and only if, $\vec{a} \in
  R^{\mathcal{B}}$. Let $f : \ind(\rho, [n]) \rightarrow \{0,1\}$. We can define
  the $\rho$-structure $\mathcal{B}_f$ over the universe $[n]$ such that for all
  $R \in \rho$, we have $\vec{a} \in R^{\mathcal{B}_f}$ if, and only if,
  $f(\vec{a}_R) = 1$. It is easy to see that the functions $\mathcal{B} \mapsto
  f_{\mathcal{B}}$ and $f \mapsto \mathcal{B}_f$ are inverse to one another, and
  so define a bijection. Moreover, we notice that for all $\sigma \in \sym_n$ we
  have that $f_{\sigma \mathcal{B}}(\vec{a}_R) = 1$ if, and only if, $\vec{a}
  \in R^{\sigma \mathcal{B}}$ if, and only if, $\sigma^{-1} \vec{a} \in
  R^{\mathcal{B}}$ if, and only if, $(f_{\mathcal{B}} \sigma^{-1})(\vec{a}_R) =
  f_{\mathcal{B}}(\sigma^{-1} \vec{a}_R)= 1$. It follows $f_{\sigma \mathcal{B}}
  = f_{\mathcal{B}} \sigma^{-1}$.

@@

  Let $\gamma_1, \gamma_2 \in [n]^{\underline{U}}$ and let $\sigma \in \sym_n$
  be such that $\gamma_1 = \sigma \gamma_2$. We have that
  \begin{align*}
    C[\gamma_1 \mathcal{A}](g_{\text{o}}) = C[\gamma_2 \mathcal{A}](g_{\text{o}}) & \Leftrightarrow  F_C(f_{\gamma_1 \mathcal{A}}) = F_C (f_{\gamma_2 \mathcal{A}}) \\
                                                                                  & \Leftrightarrow F_C(f_{\gamma_1 \mathcal{A}}) = F_C (f_{\sigma^{-1}\gamma_1 \mathcal{A}}) \\
                                                                                  & \Leftrightarrow F_C(f_{\gamma_1 \mathcal{A}}) = F_C ( f_{\gamma_1 \mathcal{A}} \sigma).
  \end{align*}

  We now combine the above observations to prove the result. Suppose $F_C$ is
  isomorphism-invariant. Let $\gamma_1, \gamma_2 \in [n]^{\underline{U}}$ and
  let $\sigma \in \sym_n$ be such that $\gamma_1 = \sigma \gamma_2$. Then, from
  isomorphism-invariance, we have $F_C(f_{\gamma_1 \mathcal{A}}) = F_C (
  f_{\gamma_1 \mathcal{A}}\sigma)$, and so $C[\gamma_1
  \mathcal{A}](g_{\text{o}}) = C[\gamma_2 \mathcal{A}](g_{\text{o}})$.

  Suppose $C$ is invariant. Fix a bijection $\gamma_1 \in [n]^{\underline{U}}$.
  Let $\sigma \in \sym_n$ and $f: \ind(\rho, [n]) \rightarrow \{0,1\}$. Let
  $\mathcal{A} = \gamma^{-1}_1 \mathcal{B}_f$ and $\gamma_2 :=
  \sigma^{-1}\gamma_1 $. Since $C$ is invariant we have $C[\gamma_1
  \mathcal{A}](g_{\text{o}}) = C[\gamma_2\mathcal{A}](g_{\text{o}})$, and so
  $F_C (f_{\gamma_1 \mathcal{A}} \sigma) = F_C(f_{\gamma_1 \mathcal{A}})$. But
  $f_{\gamma_1 \mathcal{A}} = f_{\mathcal{B}_f} = f$, and so $F_C (f \sigma) =
  F_C (f)$. It follows that $F_C$ is isomorphism-invariant.

@@

It follows that for any symmetric circuit $C$ of order $n$ there is a
homomorphism $h$ that maps $\sym_n$ to $\aut(C)$ that takes $\sigma \in \sym_n$
to an automorphism that extends it. Suppose $C$ contains no relational gates of
non-zero arity. In that case $C$ computes a constant function. As such, in this
paper we always assume a circuit contains at least one relational gate with
non-zero arity. There there exists a relational gate in $C$ with some element of
$[n]$ appearing in the tuple labelling that gate. By symmetry it follows that
all elements of $[n]$ appear in tuples labelling relational gates in $C$. Thus
no two distinct elements of $\sym_n$ agree on all input gates, and so the
homomorphism $h$ is injective.

@@

Anderson and Dawar's argument makes use of the fact that many important circuit
properties are polynomial-time decidable for circuits defined over symmetric
bases. While we should like to construct an analogous argument, we show in
Section~\ref{sec:transparency-and-unique-extension} that for circuits that
include gates computing non-symmetric functions most of these important circuit
properties are at least as hard to decide as the graph-isomorphism problem. As
such, constructing an analogous translation, as well as establishing the
numerous other needed results whose proofs rely on certain properties being
polynomial-time decidable, would be non-trivial.

In order to proceed we will restrict ourselves to circuits with certain
constraints on how those gates computing non-symmetric functions can be wired
into the circuit. We call these circuits \emph{transparent} circuits. We show in
Sections~\ref{sec:transparency-and-unique-extensions} and~\ref{} that
transparency is sufficient to allow for the polynomial-time decidability of many
important circuit properties, while still keeping the circuit model powerful
enough for us to establish a translation from formulas to families of circuits.

Before we can formally define \emph{transparency} we need to define the
\emph{syntactic-equivalence} relation on the gates of a circuit. The intuition
is that two gates in a circuit are syntactically-equivalent if the circuits
underneath them are `hereditary similar'.

@@

We could think of any gate $g$ in the circuit as computing a function that maps
a given assignment to the input gates to the evaluation of the circuit at $g$
for the given assignment. While we would like to be able to identify gates that
compute the same function in this sense, it is not hard to show that deciding
this property is $\NP$-hard. The syntactic-equivalence relation is an
easier-to-compute refinement of this relation, instead identifying pairs of
gates if the circuits underneath them are hereditary similar.

We could also define a similar equivalence relation using the automorphisms of a
circuit. In this case we identify a pair of gates $(h, h')$ in a circuit $C$ of
order $n$ if, and only if, there exists $\sigma \in \sym_n$, $\pi, \pi' \in
\aut(C)$ both extending $\sigma$, and $g$ a gate in $C$, such that $h = \pi (g)$
and $h' = \pi' (g)$. We now show this relation is a refinement of
syntactic-equivalence.

@@

We now define the \emph{unique labels} and \emph{transparency} conditions on
circuits.

@@

There is a small technical point that needs some discussion before we continue.
The definition of a circuit allows for the inclusion of a gate such that there
is no path from this gate to any of the output gates. In other words, we can be
sure, not matter the rest of the circuit, that any such gate cannot have any
effect the function computed by the circuit. We say these gates are
\emph{redundant}.

@@

It is easy to see that we can construct from a given circuit an equivalent
circuit without redundant gates by simply removing all of the redundant gates.
Let $C := \langle G, \Omega, \Sigma, \Lambda, L \rangle$ be a circuit of order
$n$ and let $g$ be a redundant circuit in $C$. It is easy to see that $g$ is not
the child of a non-redundant gate, nor is it a relational or output gate. Thus
we can define the circuit $C' := \langle G', \Omega, \restr{\Sigma}{G'},
\Lambda, \restr{L}{G'} \rangle$, where $G' \subseteq G$ is the set of all
non-redundant gates in $C$.

It is easy to see that $C'$ computes the same query as $C$, and that each $g \in
G'$ has unique labels in $C$ if, and only if, $g$ has unique labels in $C'$.
Moreover, it can be shown that if $C$ is symmetric and $g, g' \in G$ then if $g$
and $g'$ are in the same orbit then either both $g$ and $g'$ are in $G'$ or
neither are in $G'$. As such, if $C$ is symmetric then so is $C'$. We thus have
the following Lemma.

\begin{lem}
  \label{lem:redundant-gates}
  Let $C := \langle G, \Omega, \Sigma, \Lambda, L \rangle$ be a circuit of order
  $n$. Let $G'$ be the set of non-redundant gates in $C$. Let $C' := \langle G',
  \Omega, \restr{\Sigma}{G'}, \Lambda, \restr{L}{G'} \rangle$. Then (i) $C$ and
  $C'$ compute the same query, (ii) no gate in $C'$ is redundant, (iii) if $C$
  is symmetric then $C'$ is symmetric, (iv) if $C$ has unique labels then $C'$
  has unique labels, and (v) if $C$ is transparent then $C'$ is transparent
\end{lem}

We note that the construction of $C'$ from $C$ in
Lemma~\ref{lem:redundant-gates} can be implemented so as to run in time
polynomial in the size of $C$. As such, for the remainder of this paper we
assume, unless stated otherwise, every circuit has no redundant gates. With this
assumption in mind we now show that every circuit with unique labels has unique
extensions.

@@

\begin{prop}
  Let $C := \langle G, \Omega, \Sigma, \Lambda, L\rangle$ be a $(\mathbb{B},
  \rho)$-circuit of order $n$. If $C$ has unique labels then $C$ has unique
  extensions.
  \label{prop:unique-labels-unique-extensions}
\end{prop}
\begin{proof}
  Suppose $C$ has unique labels. Let $\pi_{\sigma}, \pi_{\sigma}' \in \aut(C)$
  be automorphisms extending $\sigma \in \sym_n$. We now prove that $\pi_\sigma
  = \pi_{\sigma}'$. Let $\pi := \pi_{\sigma}'\pi^{-1}_\sigma$. We have that
  $\pi$ is an automorphism of the circuit and that $\pi$ extends
  $\sigma^{-1}\sigma = e$, the identity permutation. It follows that $\pi$ fixes
  all input gates and, since for any output gate $g$, $\pi (g) = \pi
  (\Omega(\vec{a})) = \Omega(e \vec{a}) = \Omega(\vec{a}) = g$ for all $\vec{a}
  \in [n]^q$, $\pi$ fixes all output gates.

@@

We show in Proposition~\ref{prop:transparent-unique} that a transparent circuit
can be converted into an equivalent circuit with unique labels in time
polynomial in the size of the circuit. Since we are
working with $P$-uniform families of transparent circuits, this fact allows us
to assume, without a loss of generality, that transparent circuits have unique
labels (and so unique extensions).

@@

In many contexts, circuits are defined over a basis of symbols, usually the
standard logical symbols $\land, \neg, \lor$, and perhaps $\maj$ (for majority).
When evaluating a circuit, a gate labelled by a particular symbol is evaluated
using a Boolean function appropriate to the particular symbol. We have instead
defined a basis to be a set of Boolean functions. However, when reasoning about
circuits or dealing with them as inputs to algorithms we often want to identify
a function in the basis with a particular symbol. We now formally define a few
natural bases used in this paper, assigning to each function in these bases an
appropriate symbol.
